### 2025-10-06 thesis/design/functions.tex
Polish the design section

\subsubsection{Query Function – $nq$}

@@

\subsubsection{Selection Function – $\sigma$}\label{design:selection-function}

The selection function $\sigma$ retrieves the most relevant examples from the
historical embedding space $\mathcal{V}$ based on semantic similarity to the
input query. This function implements the core example selection mechanism 
that subsequently enables few-shot learning for SQL generation.

@@

\begin{enumerate}
    \item \textbf{Question similarity} – Semantic similarity between the masked input
        query and historically observed natural language queries in the embedding space
        $\mathcal{V}$, measured using cosine distance between their respective embeddings.
    \item \textbf{SQL similarity} – Structural similarity between the zero-shot inference
        result $z$ and candidate SQL queries using masked SQL representations, enabling
        pattern recognition across different database domains.
    \item \textbf{Database similarity} – Schema compatibility measured through the distance
        index $\mathcal{D}$, ensuring selected examples operate on analogous database
        structures with similar table relationships and column types.
\end{enumerate}

@@

\subsubsection{Subsetting Function – $\phi$}\label{design:subsetting-function}

@@

\subsubsection{Projection Function – $\pi$}\label{design:projection-function}

The projection function $\pi$ represents the core translation mechanism that converts
natural language queries into SQL statements using the selected examples and database
schema. This function encapsulates the large language model inference that generates
candidate SQL queries based on the provided context.

@@

\subsubsection{Refinement Function – $\rho$}\label{design:refinement-function}

@@

\subsubsection{Voting Function – $\nu$}\label{design:voting-function}

>>>>

\subsubsection{Query – $nq$}

@@

\subsubsection{Example Selection – $\sigma$}\label{design:selection-function}

The selection function $\sigma$ retrieves the most relevant examples from the
historical embedding space $v$ based on semantic similarity to the
input query. This function implements the core example selection mechanism 
that subsequently enables few-shot learning for SQL generation.

@@

\begin{enumerate}
    \item \textbf{Question similarity} – Semantic similarity between the masked input
        query and historically observed natural language queries in the embedding space
        $v$, measured using cosine distance between their respective embeddings.
    \item \textbf{SQL similarity} – Structural similarity between the zero-shot inference
        result $z$ and candidate SQL queries using masked SQL representations, enabling
        pattern recognition across different database domains.
    \item \textbf{Database similarity} – Schema compatibility measured through the distance
        index $d$, ensuring selected examples operate on analogous database
        structures with similar table relationships and column types.
\end{enumerate}

@@

\subsubsection{Schema Subsetting – $\phi$}\label{design:subsetting-function}

@@

\subsubsection{Query Projection – $\pi$}\label{design:projection-function}

The projection function $\pi$ represents the core translation mechanism that converts
natural language queries into SQL statements using the selected examples and database
schema. This function encapsulates the inference using LLMs that generates
candidate SQL queries based on the provided context.

@@

\subsubsection{Self Refinement – $\rho$}\label{design:refinement-function}

@@

\subsubsection{Voting – $\nu$}\label{design:voting-function}

### 2025-10-06 thesis/design/index.tex
Polish the design section

\begin{enumerate}
    \item \textbf{Example Selection} – Identifies semantically and structurally similar
        historical examples using masked embeddings and schema distances.
    \item \textbf{Schema Subsetting} – Reduces database schema complexity by focusing
        on relevant tables and relationships
    \item \textbf{Query Generation} – Few-shot learning with fine tuned models translate natural language queries to SQL
    \item \textbf{Self-Correction} – Iteratively refines queries through execution feedback and error analysis
    \item \textbf{Consensus Voting} – Selects the most reliable result from multiple
        generation attempts
\end{enumerate}

@@

% TODO: Make a nice diagram for this
\begin{verbatim}
  Natural Language Query 
→ Sketch Generation
→ Example Selection
→ k-times [
    → Schema Subsetting
    → Few-Shot Generation
    → Self-Correction
  ]
→ Consensus Voting
→ SQL Query
\end{verbatim}

This design builds upon concepts from DAIL-SQL \citep{DAIL} and incorporates novel
contributions in schema-aware example selection and graph-based structural similarity.

\subimport{}{initialization}

\subimport{}{components}

\subimport{}{composition}

>>>>

\begin{enumerate}
    \item \textbf{Example Selection} $\sigma$ – Identifies semantically and structurally similar examples using cosine similarity and schema distance.
    \item \textbf{Schema Subsetting} $\phi$ – Reducing schema complexity by subsetting the schema to the relevant subset of tables and relationships.
    \item \textbf{Query Projection} $\pi$ – Few-shot learning with a finetuned model to project natural language queries to SQL statements.
    \item \textbf{Self Refinement} $\rho$ – Self refinement of generated SQL statements through execution feedback and error analysis.
    \item \textbf{Voting} $\nu$ – Self consensus voting mechanism to choose the most likely result from multiple generation attempts.
\end{enumerate}

@@

% TODO: Make a nice diagram for this

\begin{verbatim}
  Natural Language Query 
→ Sketch Generation
→ Example Selection
→ k-times [
    → Schema Subsetting
    → Few-Shot Generation
    → Self-Correction
  ]
→ Consensus Voting
→ SQL Query
\end{verbatim}

This design largely builds upon in context learning concepts from DAIL-SQL \citep{DAIL-SQL}
and OmniSQL \citep{OmniSQL} but incorporates novel contributions in schema-aware example
selection by harnessing a graph-based structural similarity metric.

\subimport{}{initialization}
\subimport{}{functions}
